---
title: "Auxin Control Study September 2017 - March 2018"
author: "Bishwa"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

# Section 01
## Loading Packages

```{r setup, include= TRUE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rstatix)
library(emmeans)
library(ggpubr)
library(readxl)
```

## Read data

**auxin.data** has datashet on Sheet1, rom A2: L37. Remarks says possible 
outliers or contaminated data are; MY2:1 (row number 7), 12, and 24

```{r readata}
path<- "./Experiment_01_Auxin_Transport_Assay/auxin_tranportData.xls"

adata<- readxl::read_excel(path = path,sheet = "Sheet1",
                    range = "A1:L37",col_names = TRUE)
str(adata)

```

There are some inconsistent name variable names, lets change them.

```{r cname}

names(adata)<- c("sample", "population", "popID", "time", "hourZero", "cpm3H",
                 "perError", "dpm3H", "eff-13H", "lumex", "elapTime", "fmol"
                 )

# change character to factor

adata$population<- factor(adata$population,
                     labels =  c("Mayodan","Spiterstulen"),
                     levels =  c("MY", "SP")
                     )

# create replicate column for easyness
 ## first remove observation with high error rate

adata<- adata[-7,] # removing errenous observation

## create replicates
adata<- adata %>% group_by(population) %>% mutate(sid = 1:n())

head(adata[,c(2, 13)], 10)
tail(adata[, c(2,13)], 10)
table(adata$population)

adata<- adata %>% mutate(species = as.factor(popID))
table(adata$species)
```

# Section 02
## Summary Statistics

Now the data are cleaned, lets look at summary statistics

```{r sum, echo=FALSE}

(sum.data<- adata %>% group_by(population,species) %>% 
  summarise(mean = mean(dpm3H), sd   = sd(dpm3H),
            IQR = IQR(dpm3H),
            "%25 Q" = quantile(dpm3H, .25),
            "%50 Q" = quantile(dpm3H, .5),
            "%75 Q" = quantile(dpm3H, .75)
            )
 
)

```

# Section 03
## Check Assumptions for Parametric analysis: two groups 


```{r assumption}

# data = adata
(HistoPmani <- ggplot(data = adata,aes(x = dpm3H))+
 geom_histogram(mapping = aes(y = ..density..),
                   #breaks = (seq(-3,4, by = 1)),
                   colour = "gray30",
                   fill = "gray")+
theme_classic(10)
)


#----------- outlier
adata %>% group_by(population) %>% 
  identify_outliers(dpm3H)

# no outliers

#--------- Normality

# normality of dat can be checked with shapiro.test() test 
adata %>% group_by(population) %>% 
  shapiro_test(dpm3H)

#-------- QQ plot
adata %>% ggplot(aes(sample = dpm3H))+
  geom_qq(  geom = "point",
            position = "identity",
            distribution = stats::qnorm)+
  stat_qq_line()+
  facet_wrap(vars(population),scales = "free")+
  theme_bw(12)

#------- Homogeneity of variances: Levene's test

adata<- adata %>%  ungroup()

(adata %>%  levene_test(dpm3H~population))

# observations have equal variances

```


# Section 04
## Paired T-Test by populaton

```{r histo}
# H0:  mu(popSP): mu(popMY)
# H1:  Not so

(ptest<- adata %>% 
  t_test(dpm3H ~ population, paired = FALSE))

# the means  dpm-3h of SP  is not statistically different from mean of dpm-3h MY at 5% level

# create statistical plot 

(boxplot<- adata %>% ggplot(aes(x = population, y = dpm3H, color = population))+
  geom_boxplot(stat = "boxplot")+
  stat_summary(geom = "point", fun =  "mean", show.legend = FALSE)+
  xlab( "Population")+
  ylab(" Disintegration per minute (3H)")+
  scale_y_continuous(limits = c(0, 25000))+
  scale_color_manual("Population",values = c("Purple","Blue"))+
  theme(#panel.border = element_blank(), # will equivalent to theme_clasifc()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
)

# Add the p-value manually
(boxplot<- boxplot + stat_pvalue_manual(ptest,label = "T-test, p = {p}", y.position = 20000)+theme_classic(12))

```

# Section 04:- split by population


```{r anovaTEST}

 
#----------- outlier
adata %>% group_by(species) %>% 
  identify_outliers(dpm3H)

# one outlier

#adata<- adata[-adata$sample[adata$sample == 20],] # I don't know but using
# 21 deleted sample 22

#--------- Normality

# normality of dat can be checked with shapiro.test() test 
adata %>% group_by(population) %>% 
  shapiro_test(dpm3H)
# when observation # & removed the Mayodan species is not longer normal

#-------- QQ plot
adata %>% ggplot(aes(sample = dpm3H))+
  geom_qq(  geom = "point",
            position = "identity",
            distribution = stats::qnorm)+
  stat_qq_line()+
  facet_wrap(vars(population),scales = "free")+
  theme_bw(12)

#------- Homogeneity of variances: Levene's test

adata %>%  levene_test(dpm3H~population)

# observations have equal variances
```


# Section 05
## Anova by species/Individuals

```{r anovaSpecies}
# H0:  mu(species1)= mu(species2)  = ... = mu(species5)
# H1:  at least one mean is not equal 

# possible two way interaction

## create replicates
adata<- adata %>% group_by(population, species) %>% mutate(sp_id = 1:n()) %>% ungroup()

# adata %>% anova_test(dv = dpm3H, wid = sp_id, within = c(species))

(sp_test<- adata %>% anova_test(dpm3H ~ species,type = 3))



#------------ global anova by specis is significant


#------------------- which groups are significant??

# Pairwise comparisons
pwcs<- adata %>% tukey_hsd(dpm3H ~ species)

pwcs

pwcs<- pwcs %>% add_xy_position(x = "species")

ggboxplot(adata, x = "species", y = "dpm3H") +
  stat_pvalue_manual(pwcs, hide.ns = TRUE)

# the means  dpm-3h of SP  is not statistically different from mean of dpm-3h MY 

# create statistical plot 

( bplot_species<- adata %>% ggplot(aes(x = species, y = dpm3H, color = population))+
  geom_boxplot(stat = "boxplot", show.legend = TRUE)+
  stat_summary(geom = "point", fun =  "mean", show.legend = FALSE)+
  xlab( "Population")+
  ylab(" Disintegration per minute (3H)")+
  scale_y_continuous(limits = c(0, 25000))+
   scale_color_manual("Population",values = c("Purple","Blue"))+
  theme(#panel.border = element_blank(), # will equivalent to theme_clasifc()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))+
     stat_pvalue_manual(pwcs, hide.ns = TRUE)
)

(bplot_species<- bplot_species + stat_pvalue_manual(pwcs,label = "T-test, p.adj = {p.adj}",hide.ns = TRUE)+theme_classic())
                                                    
```

##  Anova by Interpretation by species

```{r anoReport}
#AnovaTable
sp_test %>% get_anova_table()


sp_anova<-  adata %>% 
  anova_test(
    dv        = dpm3H,
    wid       = sample,
    between   = species,
    type = 3)
sp_anova
# extract the anova and adjust p-value
main.spChange<- get_anova_table(sp_anova) %>% 
  adjust_pvalue(method = "fdr")
main.spChange
#---- pairwise comparisons
# multiple pairwise comparisons,
# usually after global anova is significant people usually use 
# tukey post-hock test which adjust p values
# however,same results canbe obtained from pair-wise_t-test, 
# don't get confused thining they are different.

# See below all approaches gives you the similar results 

# Method 1: HSD
mct<- adata %>% tukey_hsd(dpm3H ~species)

# Method2: fdr adjusted

adata %>% pairwise_t_test(formula = dpm3H ~species, p.adjust.method = "fdr")

# Method 3: bonferroni adjusted

adata %>% pairwise_t_test(formula = dpm3H ~species, p.adjust.method = "bonferroni")
# same result; see plot above


#----- 

```

## Interpretation

The anova table

`r data.frame(main.spChange)[]`


Above table indices that

there is a statistically significant difference in mean Disintegration per minute score
exists among different species (here I called species for individual in your data),
F(`r main.spChange[1,2]`, `r main.spChange[1,3]`) = `r main.spChange[1,4]`, 
p = `r main.spChange[1,8]`

Based on Table below

```{r mctTable}
mct
```

The tukey's post-hoc test shows that only  MY1 and SP1 are statistically different
(p.adj = `r mct[3,8]`)

```{r NestTed}

# #---------- aov approach
# 
aov_nest<- aov(formula = dpm3H ~ population/sp_id, data = adata)

summary(aov_nest)
#--- no interaction 

adata$sp_id<- as.factor(adata$sp_id)

 
aov_lm<- lm(dpm3H ~ population * sp_id, data = adata)

aov_lm

int_anova<- adata %>% anova_test(dpm3H ~ population * sp_id,type = 3)

int_anova

# no main and interaction effect
 
# only population  main effects  

main_pop.spp<- adata %>% group_by(population) %>% anova_test(dpm3H ~ sp_id, error = aov_lm)

 

 main_pop.spp

```



```{r finteraction}

int_anova_1<- adata %>% anova_test(dpm3H ~ population * sp_id)


(int_anova_1)

# pairwise comparisons
pwc <- adata %>% 
  group_by(population) %>%
  emmeans_test(dpm3H ~ sp_id, p.adjust.method = "bonferroni") 
pwc

# Re verification of no interaction effect


```

Interaction effect when species is nested within the population is not significant, therefore, you can rely on species effect on dpm or population effect on dpm


```{r sessioninfo}

sessionInfo()

```